% HACK: disable page numbering
% \pagenumbering{gobble}
\chapter{Diagrammer Semi-Structured Interview Protocol}
\label{app:interview-protocol}

\section{Introduction and Kick-off}
\begin{enumerate}
    \item \textit{Interviewer(s) self-introduction}
    \item Introduction to the study: Thanks again for participating in our research. We would like to know your visualization process and diagramming experience to inform our design of the next generation of visualization tools that we are currently working on.
    \item What do you mostly work on?
    \item Do you use visualization? By “visualization” we are not only referring to physical diagrams, but also to “pictures in your head.”
\end{enumerate}

\section{Past Diagramming Experience}
\begin{enumerate}
    \item What was the last visualization you made?
    \begin{enumerate}
        \item If answered with a mental visualization that only happened conceptually:
        \begin{enumerate}
            \item What did you visualize?
            \item Why didn't you draw it?
            \item What was the last diagram you made? (jump to physical visualization part)
        \end{enumerate}
        \item If answered with a physical diagram: (diagram questions)
        \begin{enumerate}
            \item What did you visualize?
            \item For what purpose did you produce the visualization?
            \item Can you tell us about the process in which you came up with and produced the diagram?
            \item What was the most challenging/time-consuming/tedious part of making this particular diagram?
            \item Can you give me a couple more examples of diagrams you've created with purpose X?
        \end{enumerate}
    \end{enumerate}
    \item Can you remember one diagram you made that you are most proud of?
    \begin{enumerate}
        \item Ask diagram questions
        \item Why are you proud of it?
    \end{enumerate}
    \item Can you remember the most challenging and/or time-consuming diagram you made?
    \begin{enumerate}
        \item Ask diagram questions (i), (ii), (iii)
        \item What made it challenging?
        \item What took so long?
    \end{enumerate}
    \item Are there diagrams that didn't make it into the final version of a particular work, such as a paper or a slide deck, you have done but were included in the drafts?
    \begin{enumerate}
        \item Why were they not included?
    \end{enumerate}
\end{enumerate}

\section{Diagramming Practice and Tools}
\begin{enumerate}
    \item What diagramming tools, libraries, and/or packages are you familiar with?
    \begin{enumerate}
        \item Which tools do you currently use? Tell me about them.
        \begin{enumerate}
            \item If multiple tools are mentioned:
            \begin{enumerate}
                \item Which one(s) do you use most frequently?
                \item In what cases is each of them used?
                \item Do you use multiple of them at a time?
            \end{enumerate}
            \item If a single tool is mentioned:
            \begin{enumerate}
                \item Why do you consistently use it?
            \end{enumerate}
        \end{enumerate}
        \item Can you tell me about the tools you know but no longer use, if any?
    \end{enumerate}
\end{enumerate}

\section{Reuse Questions}
\begin{enumerate}
    \item Introductory questions
    \begin{enumerate}
        \item Do you ever use earlier diagrams as a basis for new diagrams?
        \item Are there features of some of your diagrams that are common?
    \end{enumerate}
    \item After the introduction
    \begin{enumerate}
        \item What is the process you use to start from one diagram and move the next?
        \item What aspects of the starter diagrams are you making use of?
        \item What is the most tedious part of this process?
        \item Do you use any features of your tool to make this process easier?
        \item Are there any limitations to this tool?
    \end{enumerate}
    \item Meta-questions
    \begin{enumerate}
        \item How did you borrow feature X from diagram A to make diagram B?
        \item What tool support do you take advantage of (e.g. style libraries, templates, copy and paste of code, etc.)?
        \item When is the tool support inadequate for you to reuse your diagrams?
        \item What reuse work is particularly tedious and burdensome?
    \end{enumerate}
\end{enumerate}

\section{Audience Reception}
\begin{enumerate}
    \item How do people interact with your diagrams?
    \item What do students, readers, audience members, peers, reviewers, etc. do?
    \item Do they talk about the diagram, point to parts of it, use it in relation to some prose?
\end{enumerate}
